---
title: "Licencing Income"
author: "Bob"
date: '2022-03-02'
output:
  html_document: default
  pdf_document: default
fig_width: 6
fig_height: 4
---

In this notebook we explore how much money is made from licencing IP, particularly software licences at UK higher education providers. Academic years covered are from 2014/15 through to 2019/20. All money values are in millions of £ unless otherwise stated.

For context the University of Sheffield's research income in 2020/21 was £164m (https://www.sheffield.ac.uk/finance/annual-report-financial-statements).

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width=10, fig.height=7, dpi=300, dev="CairoPNG")
library(magrittr)
library(dplyr)
library(readr)
library(ggplot2)
library(forcats)
library(kableExtra)
library(Cairo)
```


## Source data

Source https://www.hesa.ac.uk/data-and-analysis/business-community/ip-and-startups

```{r}
table_4c <- readr::read_delim("https://www.hesa.ac.uk/data-and-analysis/providers/business-community/table-4c.csv", skip=11)
```

```{r}
table_4c <- table_4c %>%
  mutate(Value = Value / 1000)
```


# Top Ten IP Income

Average IP income over all available academic years (6 years).

```{r}
top_ten_IP <- table_4c %>%
  group_by(`HE Provider`) %>%
  summarise(`Mean IP Value (million £s)` = sum(Value) / 6) %>%
  top_n(10, `Mean IP Value (million £s)`) %>%
  mutate(`HE Provider` = fct_reorder(`HE Provider`, `Mean IP Value (million £s)`))

kbl(top_ten_IP %>% 
  arrange(-`Mean IP Value (million £s)`) %>%
  mutate(`Mean IP Value (million £s)` = round(`Mean IP Value (million £s)`, digits=1))) %>%
  kable_styling()
  
```

# All data from Top Ten IP

Time series split by IP type for top ten institutions.

```{r}
top_ten_IP_all_data <- top_ten_IP %>%
                      left_join(table_4c)
```

```{r}
top_ten_IP_all_data %>% 
  ggplot(aes(x=`Academic Year`, y=Value, fill=`Income source`)) +
    geom_col() +
    facet_wrap(~ `HE Provider`, nrow=5) +
    ylab("Value millions £")
```

# Top Ten Software Income

Average software income over all available academic years (6 years).

```{r}
top_ten_software <- table_4c %>%
  filter(`Income source` == "Software licences income") %>%
  group_by(`HE Provider`) %>%
  summarise(`Mean Software IP Value (millions £)` = sum(Value) / 6) %>%
  top_n(10, `Mean Software IP Value (millions £)`) %>%
  mutate(`HE Provider` = fct_reorder(`HE Provider`, `Mean Software IP Value (millions £)`))

kbl(top_ten_software %>% 
  arrange(-`Mean Software IP Value (millions £)`) %>%
  mutate(`Mean Software IP Value (millions £)` = round(`Mean Software IP Value (millions £)`, digits=1))) %>%
  kable_styling()
```


# All data from Top Ten software

Time series software licencing income for top ten institutions.

```{r}
top_ten_software_all_data <- top_ten_software %>%
                      left_join(table_4c)
```


```{r}
top_ten_software_all_data %>% 
  filter(`Income source` == "Software licences income") %>%
  ggplot(aes(x=`Academic Year`, y=Value)) +
    geom_col() +
    facet_wrap(~ `HE Provider`, nrow=5) +
    ylab("Value millions £")
```