---
title: "Licencing Income (WIP)"
author: "Bob"
date: '2022-03-02'
output:
  html_document: default
  pdf_document: default
fig_width: 6
fig_height: 4
---

In this notebook we explore how much money is made from licencing IP, particularly software licences at UK higher education providers. Academic years covered are from 2014/15 through to 2019/20. All money values are in millions of £ unless otherwise stated.

Source code here https://github.com/RSE-Sheffield/soft_income

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width=10, fig.height=7, dpi=300, dev="CairoPNG")
library(magrittr)
library(dplyr)
library(readr)
library(ggplot2)
library(forcats)
library(kableExtra)
library(Cairo)
```


## Source data

Sources:

* https://www.hesa.ac.uk/data-and-analysis/finances/table-1.csv
* https://www.hesa.ac.uk/data-and-analysis/business-community/ip-and-startups, 

```{r}
table_1 <- readr::read_delim("https://www.hesa.ac.uk/data-and-analysis/finances/table-1.csv", skip=12)
table_4c <- readr::read_delim("https://www.hesa.ac.uk/data-and-analysis/providers/business-community/table-4c.csv", skip=11)
```

```{r}
table_1 <- table_1 %>%
  rename(Value = `Value(£000s)`) %>%
  mutate(Value = as.numeric( gsub("\\)", "", gsub("\\(", "-", Value)))) %>%
  mutate(Value = Value / 1000)

table_4c <- table_4c %>%
  mutate(Value = Value / 1000)
```


```{r}
table_1 <- table_1 %>%
  filter(`HE Provider` != "Total")
```


# Top Ten IP Income

Average IP income over all available academic years (6 years).

```{r}
top_ten_IP <- table_4c %>%
  group_by(`HE Provider`) %>%
  summarise(`Mean IP Value (million £s)` = sum(Value) / 6) %>%
  top_n(10, `Mean IP Value (million £s)`) %>%
  mutate(`HE Provider` = fct_reorder(`HE Provider`, `Mean IP Value (million £s)`))

kbl(top_ten_IP %>% 
  arrange(-`Mean IP Value (million £s)`) %>%
  mutate(`Mean IP Value (million £s)` = round(`Mean IP Value (million £s)`, digits=1))) %>%
  kable_styling()
  
```

# All data from Top Ten IP

Time series split by IP type for top ten institutions.

```{r}
top_ten_IP_all_data <- top_ten_IP %>%
                      left_join(table_4c)
```

```{r}
top_ten_IP_all_data %>% 
  ggplot(aes(x=`Academic Year`, y=Value, fill=`Income source`)) +
    geom_col() +
    facet_wrap(~ `HE Provider`, nrow=5) +
    ylab("Value millions £")
```

# Top Ten Software Income

Average software licencing income over all available academic years (6 years).

```{r}
top_ten_software <- table_4c %>%
  filter(`Income source` == "Software licences income") %>%
  group_by(`HE Provider`) %>%
  summarise(`Mean Software IP Value (millions £)` = sum(Value) / 6) %>%
  top_n(10, `Mean Software IP Value (millions £)`) %>%
  mutate(`HE Provider` = fct_reorder(`HE Provider`, `Mean Software IP Value (millions £)`))

kbl(top_ten_software %>% 
  arrange(-`Mean Software IP Value (millions £)`) %>%
  mutate(`Mean Software IP Value (millions £)` = round(`Mean Software IP Value (millions £)`, digits=1))) %>%
  kable_styling()
```


# All data from Top Ten software

Time series software licencing income for top ten institutions.

```{r}
top_ten_software_all_data <- top_ten_software %>%
                      left_join(table_4c)
```


```{r}
top_ten_software_all_data %>% 
  filter(`Income source` == "Software licences income") %>%
  ggplot(aes(x=`Academic Year`, y=Value)) +
    geom_col() +
    facet_wrap(~ `HE Provider`, nrow=5) +
    ylab("Value millions £")
```

# Top Ten Research Income

Average reserach income over all available academic years (5 years).

```{r}
top_ten_research <- table_1 %>%
  filter(Category == "Research grants and contracts") %>%
  group_by(`HE Provider`) %>%
  summarise(`Mean Research Income (millions £)` = sum(Value) / 5) %>%
  top_n(10, `Mean Research Income (millions £)`) %>%
  mutate(`HE Provider` = fct_reorder(`HE Provider`, `Mean Research Income (millions £)`))

kbl(top_ten_research %>% 
  arrange(-`Mean Research Income (millions £)`) %>%
  mutate(`Mean Research Income (millions £)` = round(`Mean Research Income (millions £)`, digits=1))) %>%
  kable_styling()
```

# All data from Top Ten research income

Time series research income for top ten institutions.

```{r}
top_ten_research_all_data <- top_ten_research %>%
                      left_join(table_1)
```

```{r}
top_ten_research_all_data %>% 
  filter(Category == "Research grants and contracts") %>%
  ggplot(aes(x=`Academic year`, y=Value, label=round(Value, digits = 1))) +
    geom_col() +
    geom_text(size = 3, nudge_y = 5000) +
    facet_wrap(~ `HE Provider`, nrow=5) +
    ylab("Value millions £")
```
